---
title: "40_model_preparation"
output: html_notebook
---

# Prepare data for modeling 

The dataset is now ready for modeling. To make things easier in the future, I will go ahead and move the "gf" and "ga" columns towards the left side of the dataset, since they are really my target variables. 

Keep in mind, the "score" column will not necessarily match the gf and ga columns, as the former is the score from the total 60+ time on ice, and the gf and ga values are only from time when the teams each had five skaters present. 

```{r move gf and ga to left}
games <- games %>% 
  select(date, team, gf, ga, everything())
```

Finally, let's take another look at the data. 
```{r look at data again}
head(games)
```

There are many options for which variables one would include in a model. A few examples: 

1. The effect of shot quality (high-, mid-, or low-danger) in predicting goals for and against each team. 

2. Is PDO (shooting percentage + save percentage and a common measure of "luck") indeed the most important factor in winning a game?

3. The effect (if any) of things like attendance and time of year on on-ice performance. 
